How to write a PhD in Biological Sciences: a guide for the uninitiated

### 2021-07-20 3.04-MaterialsMethods.Rmd
Add files via upload
# After copy edits

## The Materials and Methods {#methods}

@@

### Sections or subheadings

As ever, I’m going to recommend inserting section headers into your Materials and Methods to help you break it up into bite sized chunks. These are principally for the reader, but will also make your job easier.

An important point here is to make sure that you keep your variables (what you are measuring to test your question) out of the sections like study organism and animal husbandry. Don’t be tempted to scatter variables and how you measured them throughout the Materials and Methods section. Keep them clear and maintain a logical flow, using a time-line if that makes sense.

### Study organism or study system

Having already written the [introduction](introduction.html), you will know that there isn’t too much space to write a lot of information about your study organism there. However, often there is important background information that is needed on the species, which would be distracting or lose the flow of the [introduction](introduction.html). By having a ‘Study Organism’ section at the start of your Materials and Methods, you can add in all relevant information in a paragraph or two.

For example, if you are working on African clawed frogs, _Xenopus laevis_, as a model organism you will want to use this space to describe why they are a good system for this particular investigation. Alternatively, your study could be detailing their invasion biology, in which case this would be the section where you can explain their distribution, that they are principally aquatic, or that they are invasive on four continents. You probably won’t need all of these points, but just the aspects of this species’ biology that is relevant to your study. You should make sure you know whether or not the journal requires the taxonomic authority. If you are not sure, include it now ([see here for an explanation](taxonomy.html)).

Similarly, if you are working on invasive zebra fish, _Danio rerio_, you can write in this section the known dates when they were introduced to Japan, and how we know that those animals came from. You can also give details of their natural range in South Asia.

@@

### Animal husbandry / Growing Conditions / Culturing Techniques
If you have kept animals in the lab, grown plants or cultured bacteria in order to do your experiments, then you need to provide details of how you did this. For plants this will include all relevant information on their source, propogation, growth medium and climatic regimes. For animals, you need to include their source, welfare, housing, feeding, light:dark cycles, temperature, etc. If you bred animals to produce your study life-history stage, all relevant information is needed here. All this information should be available from your ethics application. Some journals will want you to include the details of your ethics permission here.

### Repetitive methodology
If you’ve used the same methodology for each part of your data collection (like DNA extraction, sequencing, etc.), you should have a section that explains all of this prior to any experimental manipulations done. I would suggest that this is detailed in the first relevant chapter of your thesis, and then referred to in later chapters. It’s best to go to look to see what other people have reported on in order to know what level of detail to include. For example, 20 years ago it would have been important to include detailed information of how you extracted DNA from tissues. Today, you can probably say that you used a standard extraction kit or salt extraction without saying more.

### Do you need to include formulae?
Some of your methodologies will include calculations that rest on mathematical formula which can be written in mathematical notation. Certainly, there are plenty of examples in the biological sciences, especially in modelling, where these formulas are needed because they are developed within the methodology in order to produce the [results](results.html). If your thesis involves this type of mathematical progression, then you will already know this is expected because all of the literature that you read will contain equations and formulas. However, if you are calculating (for example) the surface areas of a leaf, through a set of measurements, do you need to provide the formula you used in your materials and methods?

@@

### Experimental manipulations and data collection
If you have done more than one kind of experiment, it’s probably a good idea to give each a separate subheading. Be consistent with these subheadings between the Materials and Methods and the  [results](results.html).

One very important point here is to fully explain the collection of all of the variables that are used in your analyses. Remember to use exactly the same names for your variables here, as you have in the [introduction](introduction.html) (and will use in your [figures](figures.html) and  [results](results.html) sections). It’s very important that the reader gets this consistency across sections. Please also remember to include all units that you collect data in (there is a space between numbers and units!), and the accuracy of the measuring equipment used.

@@

#### Equipment

Many journals require that you name the company that you bought equipment from as well as the town and country where it was made. This is to help others that might want to buy the same equipment (although it’s a bit outdated in these days of multinational companies). You should try to get all of the model numbers of equipment as you do the work, so that you are not scrambling later on to find out what they were. It’s really quick and easy to take a picture of this at the time.

@@

### Data analysis
This section is getting increasingly longer as people do ever more fancy statistical manipulations. As above, it’s very important that you use the same names for the variables as you have introduced and collected them. In addition to explaining the specific tests that you performed, you should explain what roles the variables have as dependent, independent, or random variables in your model.

@@

You should be familiar with the terms above from when you formulated your hypothesis (see [Part II](hypothesis.html) for a refresher).

@@

Give the name for each package in R that you use, together with its citation. Some journals will have particular software conventions that you need to respect and adhere to. Also give a citation to R as well as the version that you used in your analysis. If you don’t use the latest version of R, then your reviewers may want to know why, so always remember to update your version of R when you do your analysis.

### Did you leave anything out?
By following your Materials and Methods, anyone else should be able to repeat your work. If there isn’t enough detail for someone else to conduct the same experiment or survey, then you will need to add this information somewhere. For example, a list of sampling points or a database of sequences might be needed in Supplementary Information.

>>>>

# The Materials and Methods {#methods}

@@

## Sections or subheadings

As ever, I’m going to recommend inserting section headers into your Materials and Methods to help you break it up into bite sized chunks (Figure \@ref(fig:methods)). These are principally for the reader, but will also make your job easier.

An important point here is to make sure that you keep your variables (what you are measuring to test your question) out of the sections like study organism and animal husbandry. Don’t be tempted to scatter variables and how you measured them throughout the Materials and Methods section. Keep them clear and maintain a logical flow, using a timeline if that makes sense.

## Study organism or study system

Having already written the [introduction](#introduction), you will know that there isn’t too much space to write a lot of information about your study organism there. However, often there is important background information that is needed on the species, which would be distracting or lose the flow of the [introduction](#introduction). By having a ‘Study Organism’ section at the start of your Materials and Methods, you can add in all relevant information in a paragraph or two.

For example, if you are working on mice, _Mus muculus_, as a model organism you will want to use this space to describe why they are a good system for this particular investigation. Alternatively, your study could be detailing their invasion biology, in which case this would be the section where you can explain their distribution, or that they are invasive on all continents. You probably won’t need all of these points, but just the aspects of this species’ biology that is relevant to your study. You should make sure you know whether or not the journal requires the taxonomic authority. If you are not sure, include it now ([see here for an explanation](#taxonomy)).

Similarly, if you are working on invasive zebra fish, _Danio rerio_, you can write in this section the known dates when they were introduced to Japan, and how we know where those animals came from. You can also give details of their natural range in South Asia.

@@

## Animal husbandry / Growing conditions / Culturing techniques
If you have kept animals in the lab, grown plants or cultured bacteria in order to do your experiments, then you need to provide details of how you did this. For plants, this will include all relevant information on their source, propagation, growth medium and climatic regimes. For animals, you need to include their source, welfare, housing, feeding, light:dark cycles, temperature, etc. If you bred animals to produce your study life-history stage, all relevant information is needed here. All this information should be available from your ethics application. Some journals will want you to include the details of your ethics permission here.

## Repetitive methodology
If you’ve used the same methodology for each part of your data collection (like DNA extraction, sequencing, etc.), you should have a section that explains all of this prior to any experimental manipulations done. I would suggest that this is detailed in the first relevant chapter of your thesis, and then referred to in later chapters ([see section on generative recycling](#generative)). It’s best to go to look to see what other people have reported on in order to know what level of detail to include. For example, 20 years ago, it would have been important to include detailed information of how you extracted DNA from tissues. Today, you can probably say that you used a standard extraction kit without saying more.

## Do you need to include formulae?
Some of your methodologies will include calculations that rest on mathematical formula which can be written in mathematical notation. Certainly, there are plenty of examples in the biological sciences, especially in modelling, where these formulas are needed because they are developed within the methodology in order to produce the [results](#results). If your thesis involves this type of mathematical progression, then you will already know this is expected because all of the literature that you read will contain equations and formulas. However, if you are calculating (for example) the surface areas of a leaf, through a set of measurements, do you need to provide the formula you used in your materials and methods?

@@

## Experimental manipulations and data collection
If you have done more than one kind of experiment, it’s probably a good idea to give each a separate subheading. Be consistent with these subheadings between the Materials and Methods and the  [results](#results).

One very important point here is to fully explain the collection of all of the variables that are used in your analyses. Remember to use exactly the same names for your variables here, as you have in the [introduction](#introduction) (and will use in your [figures](#figures) and  [results](#results) sections). It’s very important that the reader gets this consistency across sections. Please also remember to include all units that you collect data in (there is a space between numbers and units!), and the accuracy of the measuring equipment used.

@@

### Equipment

Many journals require that you name the company that you bought equipment from as well as the town and country where it was made. This is to help others that might want to buy the same equipment (although it’s a bit outdated in these days of multinational companies). You should try to get all of the model numbers of equipment as you do the work, so that you are not scrambling, later on, to find out what they were. It’s really quick and easy to take a picture of this at the time.

@@

## Data analysis
This section is getting increasingly longer as people do ever more intricate statistical manipulations. As above, it’s very important that you use the same names for the variables as you have introduced and collected them. In addition to explaining the specific tests that you performed, you should explain what roles the variables have as dependent, independent, or random variables in your model.

@@

You should be familiar with the terms above from when you formulated your hypothesis (see [Part II](#hypothesis) for a refresher).

@@

Give the name for each package in R that you use, together with its citation. Some journals will have particular software conventions that you need to respect and adhere to. Also, give a citation to R as well as the version that you used in your analysis. If you don’t use the latest version of R, then your reviewers may want to know why, so always remember to update your version of R when you do your analysis.

## Did you leave anything out?
By following your Materials and Methods, anyone else should be able to repeat your work. If there isn’t enough detail for someone else to conduct the same experiment or survey, then you will need to add this information somewhere. For example, a list of sampling points or a database of sequences might be needed in Supplementary Information.

### 2021-07-20 3.05-Results.Rmd
Add files via upload
# After copy edits

## The correct approach - responding to your hypotheses
Yes, the easiest way to write the results is simply to have the questions or hypotheses that you posed at the end of the introduction and have them as subheadings for your results section. Once you have these subheadings, you simply assemble all of the results with respect to that hypothesis. You should already have all the tests that you used in the methods section.

@@

When describing the relationship between two variables, make sure that you state the direction of the effect. This means not simply stating that they have a significant relationship, but what that relationship is. This means stating that one increases with the other (a direct relationship) or one increases when the other decreases (an inverse relationship). Once you have described the result, either make reference to the Table or Figure where this can be seen, or describe the results of the test in brackets.

@@

(ref:results) **You may well find that your results section is the shortest section of your manuscript.** Only report what you need to in this section. Keep results that don't respond to your hypotheses for the Supplementary Information.
```{r results, echo=FALSE, out.width = '95%', fig.cap="(ref:results)"}
knitr::include_graphics('figures/results.png')
```

@@

## Descriptive results
Some studies have a set of descriptive results that, by convention, are placed without a section header at the start of the results. These normally convey the size of the dataset that entered into the rest of the analyses. If the size of the dataset varies by design in the methods, then you don’t need it here. But you can’t always control the sample size, so it needs to be established before moving on to the rest of the results. In genetic studies, for example, this might include the number of samples that were collected and successfully sequenced, and thus entered into the rest of the analyses. Your text can then be broken into sections that correspond with your questions, unless you only have one in which case further sections are not warranted.

## Responding to hypotheses
While I encourage you to present data that corresponds to hypotheses here in the results section, you should not formally accept or reject your hypothesis here. This is done in the discussion, and allows you to present caveats and other important issues.

## Common Mistakes
- Having extra analyses in the results that you didn’t include in your methods. If you really didn’t anticipate conducting the extra tests, then these are actually _post hoc_ (after the event) tests (not anticipated before the study) and you should think carefully about how to treat these. Otherwise, you’ll need to r-ejig your methods to make sure that all tests are included.
- Meaningless precision. This can happen throughout a thesis or manuscript (e.g. in the Materials and Methods), but it is most common in the results when too many significant figures are reported. For example, the output of your analysis might tell you that flight feathers in one population are 2.6484332 mm longer than those of another (± 0.0654562 mm), but in your methods you only use callipers that have a precision of 0.01 mm. Reporting results beyond the precision whereby they can be measured quickly becomes meaningless. Whenever you obtain the output of a statistical package, you need to make sure that you use your practical head to re-interpret these results and think about what they actually mean. 
- Repeating numbers in the text that are already in tables or graphs. The text is likely to be packed with numbers already, so don’t repeat any that are already represented elsewhere.
- Repeating text in different sections. As you’ve seen elsewhere in this book, repeating text simply isn’t acceptable. It doesn’t matter if you did the same thing many times, you can always write something in a different way. Having repeated text is the fastest way to bore the pants off your reader, and therefore must be avoided at all costs.

>>>>

## The correct approach - responding to your hypotheses
Yes, the easiest way to write the results is simply to have the questions or hypotheses that you posed at the end of the introduction and have them as subheadings for your results section (Figure \@ref(fig:results)). Once you have these subheadings, you simply assemble all of the results with respect to that hypothesis. You should already have all the tests that you used in the methods section.

@@

When describing the relationship between two variables, make sure that you state the direction of the effect. This means not simply stating that they have a significant relationship, but what that relationship is. This means stating that one increases with the other (a direct relationship) or one increases when the other decreases (an inverse relationship). Note that other types of variable relationships exist. Once you have described the result, either make reference to the Table or Figure where this can be seen, or describe the results of the test in brackets.

@@

(ref:results) **You may well find that your results section is the shortest section of your manuscript.** Only report what you need to in this section. Keep results that don't respond to your hypotheses for the Supplementary Information.

```{r results, echo=FALSE, out.width = '95%', fig.cap="(ref:results)"}
knitr::include_graphics('figures/results.png')
```

@@

## Descriptive results
Some studies have a set of descriptive results that, by convention, are placed without a section header at the start of the results. These normally convey the size of the dataset that entered into the rest of the analyses. If the size of the dataset varies by design in the methods, then you don’t need it here. But you can’t always control the sample size, so it needs to be established before moving on to the rest of the results. In genetic studies, for example, this might include the number of samples that were collected and successfully sequenced, and thus entered into the rest of the analyses. Your text can then be broken into sections that correspond with your questions, unless you only have one, in which case further sections are not warranted.

## Responding to hypotheses
While I encourage you to present data that corresponds to hypotheses here in the results section, you should not formally accept or reject your hypothesis in the results. This is done in the discussion, and allows you to present caveats and other important issues.

## Common mistakes
- Having extra analyses in the results that you didn’t include in your methods. If you really didn’t anticipate conducting the extra tests, then these are actually _post hoc_ (after the event) tests (not anticipated before the study) and you should think carefully about how to treat these. Otherwise, you’ll need to rejig your methods to make sure that all tests are included.
- Meaningless precision. This can happen throughout a thesis or manuscript (e.g. in the Materials and Methods), but it is most common in the results when too many significant figures are reported. For example, the output of your analysis might tell you that flight feathers in one population are 2.6484332 mm longer than those of another (± 0.0654562 mm), but in your methods you only use callipers that have a precision of 0.01 mm. Reporting results beyond the precision whereby they can be measured becomes meaningless. Whenever you obtain the output of a statistical package, you need to make sure that you use your practical head to re-interpret these results and think about what they actually mean. 
- Repeating numbers in the text that are already in tables or graphs. The text is likely to be packed with numbers already, so don’t repeat any that are already represented elsewhere.
- Repeating text in different sections. As you’ve seen [elsewhere in this book](#plagiarise), repeating text simply isn’t acceptable. It doesn’t matter if you did the same thing many times, you can always write something in a different way. Having repeated text is the fastest way to bore the pants off your reader, and therefore must be avoided at all costs.

### 2021-07-20 3.06-Discussion.Rmd
Add files via upload
# After copy edits

The discussion should be the final section of your manuscript or chapter that you are going to write (given that the [abstract](abstract.html) will come last). This is because for the discussion you must already know all of the rest of the manuscript, starting with the hypothesis, which dictates what will go into all of the other sections. While writing the other sections, I often make notes under the heading ‘discussion’ so that it acts as an _aide memoire_ to ideas that I’ve had during the study.

Before you start writing your discussion, make a plan and then discuss this plan with your advisor. I’d make this suggestion for all of the sections of your chapter or paper. It doesn’t take that long to do, and it provides an opportunity for you and your advisor to talk about the [results](results.html) of your work and discuss them together. Such discussions should be eye opening for both of you, and they provide a great opportunity for you and your advisor to get excited about the work you’ve done, your [results](results.html) and what they mean. I get a lot of enjoyment during these discussions, especially when sharing the excitement of the [results](results.html). Sharing thoughts before you start writing is important because by talking about it, you and your advisor are more likely to come to a consensus about the best way in which to interpret your [results](results.html). Conversely, presenting them with the discussion finished might not be the best way of pitching ideas to them. Once you've had time to discuss it, you might have a lot more work to do in rewriting what you've already produced.

There is a convention that you do not refer to the [tables](tables.html) or [figures](figures.html) in the discussion. This is because they have already been referred to in the [results](results.html), and a reader should have already consulted them there. Occasionally, you may want to have a new conceptual diagram in the discussion, but this is rare. Most often you do not refer to figures or tables and no new ones appear in the discussion.

In general, the first and last paragraphs of the discussion are key to the reader, but the discussion must also consider caveats and limitations in the experimental design and interpretation of your [results](results.html), as well as providing a concise discussion of the results in the context of existing literature. This is also your opportunity to suggest new hypotheses and how they could be tested.

@@

## First paragraph of the discussion
Your discussion begins by you responding to your hypothesis, clearly stating whether or not it has been accepted, and putting this into the wider context of the study (i.e. paragraph one or two of the [introduction](introduction.html) with relevant literature). You can then follow these statements by emphasising what you consider to be the most important finding, and explain how it adds to existing literature. However, don’t be tempted to over-interpret your results, or claim that they mean more than they do ([see section on speculation](speculation.html) below).

@@

## To sub-section or not to sub-section the discussion?
My preference is to plan the discussion before you write it, just as you did for the [introduction](introduction.html) using a bullet pointed list. This will provide you with logical sub-section headers for the discussion under which you can write the first draft. When your chapter has a simple aim that is easily communicated, I’d suggest deleting these sub-section headings before you finish. However, many studies are more complex and contain multiple experiments or evidential approaches. It is then sometimes wise to leave sub-sections in your discussion so that your reader can more easily follow the text. Where possible, these should be the same sub-sections that you have broken your [methods](methods.html) and [results](results.html) into, especially where these relate to specific hypotheses or aims. Or it may be more appropriate to discuss the different approaches separately, specifically when the literature that you refer to falls into different groups.

@@

The discussion is about explaining the meaning of your [results](results.html) to the reader. I often find that people write a lot of inappropriate information in the discussion. Remember that this section is not going to provide background information, and is unlikely to bring up new topics that need introducing. It may be that your [results](results.html) prompt you to introduce a new area of research that wasn’t covered in your [introduction](introduction.html), and this is fine. But for the main part, you should discuss your [results](results.html) in the context of existing literature. You can expect that the literature that you use in your discussion will only partially overlap with your [introduction](introduction.html), with plenty of new citations. Similarly, it can be that discussing your [results](results.html) will mean that you end up with paragraphs that have no citations.

In order to provide a critical evaluation of your research findings, you really need to present a reasoned argument [@jenicek2006how]. In some respects, when writing the discussion your walk a tightrope between inferring the meaning of your results for a wider audience, and making claims that are too broad (Table \@ref(tab:discuss)). Overreaching in the discussion (and consequently [abstract](abstract.html)) is a common problem that examiners and reviewers regularly ask authors pull back on their claims.

When providing different sides of an [argument](argument.html), try to use your [results](results.html) to conclude that one side is supported more than the other. If your results don't help with this particular point, then it could be that you are trying to discuss something that isn't directly related to the work. This is a very common problem in discussions, and a good test is asking yourself how your [results](results.html) add to the point you are trying to discuss. If they don't leave it out and move on.

@@

| **Do** | **Don't** |
|----|----|
| Provide a summary of your major findings | Use statements that are too broad |
| Interpret your results with respect to mechanisms | Over-represent your results |
| Give problems with methods and techniques | Provide unrealistic solutions to deficiencies |
| Compare similarities and contrasts with other studies | Ignore conflicting literature |
| Provide practical implications of your results | Be unrealistic about what could be done |
| Give directions for future research | Simply state what you have done elsewhere in your thesis |
| Restrain your conclusions to evidence from your data | Make unsubstantiated claims that overreach your evidence |
| Provide insight into unresolved questions, controversies and biases | [_Reductio ad absurdum_](argument.html)  |
| Explain the study design's limitations and uncertainties | Inflate the importance of your findings |
| Highlight the strengths of your study | Forget to mention the limitations |
| Provide alternative interpretations and hypotheses | Go into tangential issues |
| Give unexpected results and anomalies | Avoid being critical of how your study could have produced unexpected results |
| Mention literature that contradicts your findings | Use your conclusions as a 'bully pulpit' |
| Provide arguments that are logical and coherent | Erect a ['straw man' or 'false dichotomy'](argument.html) |
| Make conclusions based on your logic | Make emotional appeals to the reader |
| Make sure that you've discussed all results | Suddenly introduce new data or results not previously presented |
| Boast about your findings | Apologise or undermine your own work |
| Be concise | Waffle on |
| Take a look at target journal requirements and suggestions for a  discussion | Feel that journal advice is not relevant to you |

## Caveats and limitations
An important aspect of the discussion is to consider how the interpretation of your [results](results.html) may be incorrect. For example, if you have done an experiment, how well controlled was it and how well could it be considered to scale up to real world interactions? Could you have measured other variables? Almost every study will have caveats and limitations, and it is very important that you report them in a considered approach. You can also add ideas on how to improve your approach to eliminate important caveats that you have identified, even if this means suggesting something that adds evidence from a completely different field of biology.

@@

## Should you speculate in the discussion? {#speculation}
Reviewers will often be unhappy with speculation in the discussion section. Speculation isn’t that hard to spot, as it occurs when you make claims for which your [results](results.html) have no foundation. I think that it is healthy to have one or two statements that are speculative, but clearly labelled as such. After all, after writing this paper, you are going to be one of the world experts in the topic, and thus your deeper understanding is often worth relating to the reader. However, I suggest that you speculate in combination with suggesting what work could be done in future. If you really feel that the point has to be made, you must clearly label it as speculative.

Perhaps an easier trap to fall into is over-interpretation. This is when you suggest that your [results](results.html) mean more than they do. It’s an easy trap to fall into, especially after setting up the study in relation to key topics in the discipline (presented in paragraph one or two of the [introduction](introduction.html)). You will probably find it hard to see where you have over-interpreted, and this is something that having your work read by your advisor, or another colleague, will really help. You may then be asked to ‘tone down’ your claim, or to place it into the direct findings of your [results](results.html).

@@

## Don’t beat up on others
Your [results](results.html) may show that other researchers were wrong with their interpretation or findings. Whatever you may think of them, never use your discussion to be disrespectful to other researchers or their work. This has been referred to as the “bully pulpit”. As with all aspects of professional interactions, consider how you would like to be treated, and act accordingly. This is not to say that you shouldn’t point out mistakes that were made before, but be sure not to become emotive or insulting.

Things brings to mind Freud's narcissism of minor differences, in which he pointed out that bitter fights could start between groups when they in fact agree on 90% on issues. So it is likely to be true that someone you might disagree with one one point, is likely to be allied with you on nearly all of their other views and opinions. Making and maintaining allegiances is more likely to get you in a position where you can constructively discuss any minor disagreements. Consider this before you are tempted to launch a negative sounding social media post.

@@

## Where next?
The ‘where next’ aspect of your discussion is important as it may provide the reader with ideas for their own work. Of course, these are questions that you may wish to pursue in your own career, or they may be corroborating evidence from other disciplines that you will never undertake. Either way, making pointers for continuing aspects of the research is an important component of the discussion. Providing new lines of research may also allow you to speculate about what you consider to be the most important angle of this topic now that you have presented your [results](results.html). This should be justifiable and not gratuitous. There's no point in suggesting what other studies could be done just for the sake of it, or just because you happen to have already done them in your thesis. But if there are legitimate parts of your thesis that link to your discussion of your chapter (and this is quite likely), then you should definitely point to them.

## Last paragraph
The last paragraph of the discussion is your take home message. It’s a summary paragraph that sets out what you aimed to achieve, and what the new state of understanding of the topic is now that your [results](results.html) are out. This should include the key literature that can now be reconsidered.

>>>>

The discussion should be the final section of your manuscript or chapter that you are going to write (given that the [abstract](#abstract) will come last). This is because for the discussion you must already know all of the rest of the manuscript, starting with the hypothesis, which dictates what will go into all of the other sections. While writing the other sections, I often make notes under the heading ‘discussion’ so that it acts as an _aide memoire_ to ideas that I’ve had during the study.

Before you start writing your discussion, make a plan and then discuss this plan with your advisor. I’d make this suggestion for all of the sections of your chapter or paper. It doesn’t take that long to do, and it provides an opportunity for you and your advisor to talk about the [results](#results) of your work and discuss them together. Such discussions should be eye opening for both of you, and they provide a great opportunity for you and your advisor to get excited about the work you’ve done, your [results](#results) and what they mean. I get a lot of enjoyment during these discussions, especially when sharing the excitement of the [results](#results). Sharing thoughts before you start writing is important because by talking about it, you and your advisor are more likely to come to a consensus about the best way in which to interpret your [results](#results). Conversely, presenting them with the discussion finished might not be the best way of pitching your ideas.

There is a convention that you do not refer to the [tables](#tables) or [figures](#figures) in the discussion. This is because they have already been referred to in the [results](#results), and a reader should have already consulted them there. Occasionally, you may want to have a new conceptual diagram in the discussion, but this is rare. Most often, you do not refer to figures or tables and no new ones appear in the discussion.

In general, the first and last paragraphs of the discussion are key to the reader, but the discussion must also consider caveats and limitations in the experimental design and interpretation of your [results](#results), as well as providing a concise discussion of the results in the context of existing literature (Figure \@ref(fig:discussion)). This is also your opportunity to suggest new hypotheses and how they could be tested.

@@

## First paragraph of the discussion
Your discussion begins by you responding to your hypothesis, clearly stating whether or not you accept it, and putting this into the wider context of the study (i.e. paragraph one or two of the [introduction](#introduction) with relevant literature). You can then follow these statements by emphasising what you consider to be the most important finding, and explain how it adds to existing knowledge. However, don’t be tempted to over-interpret your results, or claim that they mean more than they do ([see section on speculation](#speculation) below).

@@

## To sub-section or not to sub-section the discussion?
My preference is to plan the discussion before you write it, just as you did for the [introduction](#introduction) using a bullet-pointed list. This will provide you with logical sub-section headers for the discussion under which you can write the first draft. When your chapter has a simple aim that is easily communicated, I’d suggest deleting these sub-section headings before you finish. However, many studies are more complex and contain multiple experiments or evidential approaches. It is then sometimes wise to leave sub-sections in your discussion so that your reader can more easily follow the text. Where possible, these should be the same sub-sections that you have broken your [methods](#methods) and [results](#results) into, especially where these relate to specific hypotheses or aims. Or it may be more appropriate to discuss the different approaches separately, specifically when the literature that you refer to falls into different groups.

@@

The discussion is about explaining the meaning of your [results](#results) to the reader. I often find that people write a lot of inappropriate information in the discussion. Remember that this section is not going to provide background information, and is unlikely to bring up new topics that need introducing. It may be that your [results](#results) prompt you to introduce a new area of research that wasn’t covered in your [introduction](#introduction), and this is fine. But for the main part, you should discuss your [results](#results) in the context of existing literature. You can expect that the literature that you use in your discussion will only partially overlap with your [introduction](#introduction), with plenty of new citations. Similarly, it can be that discussing your [results](#results) will mean that you end up with paragraphs that have no citations.

In order to provide a critical evaluation of your research findings, you really need to present a reasoned argument [@jenicek2006how]. In some respects, when writing the discussion your walk a tightrope between inferring the meaning of your results for a wider audience, and making claims that are too broad (Table \@ref(tab:discuss)). Overreaching in the discussion (and consequently [abstract](#abstract)) is a common problem that examiners and reviewers regularly ask authors pull back on their claims.

When providing different sides of an [argument](#argument), try to use your [results](#results) to conclude that one side is supported more than the other. If your results don't help with this particular point, then it could be that you are trying to discuss something that isn't directly related to the work. This is a very common problem in discussions, and a good test is asking yourself how your [results](#results) add to the point you are trying to discuss. If they don't leave it out and move on.

@@

| **Do** | **Don't** |
|----|----|
| Provide a summary of your major findings | Use statements that are too broad |
| Interpret your results with respect to mechanisms | Over-represent your results |
| Give problems with methods and techniques | Provide unrealistic solutions to deficiencies |
| Compare similarities and contrasts with other studies | Ignore conflicting literature |
| Provide practical implications of your results | Be unrealistic about what could be done |
| Give directions for future research | Simply state what you have done elsewhere in your thesis |
| Restrain your conclusions to evidence from your data | Make unsubstantiated claims that overreach your evidence |
| Provide insight into unresolved questions, controversies and biases | [_Reductio ad absurdum_](#argument)  |
| Explain the study design's limitations and uncertainties | Inflate the importance of your findings |
| Highlight the strengths of your study | Forget to mention the limitations |
| Provide alternative interpretations and hypotheses | Go into tangential issues |
| Give unexpected results and anomalies | Avoid being critical of how your study could have produced unexpected results |
| Mention literature that contradicts your findings | Use your conclusions as a 'bully pulpit' |
| Provide arguments that are logical and coherent | Erect a ['straw man' or 'false dichotomy'](#argument) |
| Make conclusions based on your logic | Make emotional appeals to the reader |
| Make sure that you've discussed all results | Suddenly introduce new data or results not previously presented |
| Boast about your findings | Apologise or undermine your own work |
| Be concise | Waffle on |
| Take a look at target journal requirements and suggestions for a  discussion | Feel that journal advice is not relevant to you |

## Caveats and limitations
An important aspect of the discussion is to consider how the interpretation of your [results](#results) may be incorrect. For example, if you have done an experiment, how well controlled was it and how well could it be considered to scale up to real-world interactions? Could you have measured other variables? Almost every study will have caveats and limitations, and it is very important that you report them in a considered approach. You can also add ideas on how to improve your approach to eliminate important caveats that you have identified, even if this means suggesting something that adds evidence from a completely different field of biology.

@@

## Should you speculate in the discussion? {#speculation}
Reviewers will often be unhappy with speculation in the discussion section. Speculation isn’t that hard to spot, as it occurs when you make claims for which your [results](#results) have no foundation. I think that it is healthy to have one or two statements that are speculative, but clearly label them as such. After all, after writing this paper, you are going to be one of the world experts in the topic, and thus your deeper understanding is often worth relating to the reader. However, I suggest that you speculate in combination with suggesting what work could be done in future. Remember, if you really feel that the point has to be made, you must clearly label it as speculative.

Perhaps an easier trap to fall into is over-interpretation. This is when you suggest that your [results](#results) mean more than they do. It’s an easy trap to fall into, especially after setting up the study in relation to key topics in the discipline (presented in paragraph one or two of the [introduction](#introduction)). You will probably find it hard to see where you have over-interpreted, and this is something that having your work read by your advisor, or another colleague, will really help. You may then be asked to ‘tone down’ your claim, or to place it into the direct findings of your [results](#results).

@@

## Don’t beat up on others
Your [results](#results) may show that other researchers were wrong with their interpretation or findings. Whatever you may think of them, never use your discussion to be disrespectful to other researchers or their work. This has been referred to as the 'bully pulpit' or an [_ad hominem_ attack](#argument). As with all aspects of professional interactions, consider how you would like to be treated, and act accordingly. This is not to say that you shouldn’t point out mistakes that were made before, but be sure not to be emotive or insulting.

Things brings to mind Freud's narcissism of minor differences, in which he pointed out that bitter fights could start between groups when they, in fact, agree on 90% on issues. So it is likely to be true that someone you might disagree with one one point, is likely to be allied with you on nearly all of their other views and opinions. Making and maintaining allegiances is more likely to get you in a position where you can constructively discuss any minor disagreements. Consider this before you are tempted to launch a negative sounding social media post.

@@

## Where next?
The ‘where next’ aspect of your discussion is important as it may provide the reader with ideas for their own work. Of course, these are questions that you may wish to pursue in your own career, or they may require corroborating evidence from other disciplines that you will never undertake. Either way, making pointers for continuing aspects of the research is an important component of the discussion. Providing new lines of research may also allow you to speculate about what you consider to be the most important angle of this topic now that you have presented your [results](#results). This should be justifiable and not gratuitous. There's no point in suggesting what other studies could be done just for the sake of it, or just because you happen to have already done them in your thesis. But if there are legitimate parts of your thesis that link to your discussion of your chapter (and this is quite likely), then you should definitely point to them.

## Last paragraph
The last paragraph of the discussion is your take-home message. It’s a summary paragraph that sets out what you aimed to achieve, and what the new state of understanding of the topic is now that your [results](#results) are out. This should include the key literature that can now be reconsidered.

### 2021-07-20 3.07-Acknowledgements.Rmd
Add files via upload
# After copy edits

## The Acknowledgements {#acknowledgements}

@@

### Thank your funders

Any people or organisations that funded you or any component of your work. Check with them they may have specific wording that they want you to use. Remember that your institution may not have given you money but may have supported you in other ways you should probably thank them. There are registers of funders, and you should use to see whether your funder is present and exactly how they should be acknowledged. One example is the [Crossref Funder Registry](https://www.crossref.org/services/funder-registry/). Funders rely on being able to search [databases](databases.html) for their name and find all of the work that they have funded. This means that you should use the particular wording that they supplied, especially if they specify this in their funding letter.

@@

### People and organisations to thank

 - Try to remember all of those people who helped you doing fieldwork, or in the laboratory. Especially laboratory technicians without whom a lot of the work that we do isn’t possible.
 - People that may have helped administer your funds or other office administration. 
 - Organisations that gave you permission to do your work. In some journals the acknowledgements is the correct place to put your permit numbers along with the names of the organisations who provided them. 
 - Increasingly, permission from ethics committees is placed in the materials and methods but do check with the journal instructions to authors.
 - People who supplied you with samples, images or other elements of your work or presentation of your manuscript.

@@

### Who not to thank (in your manuscript acknowledgements)
 - You do not need to thank anyone who is an author. The exception is in your thesis acknowledgements, where you should mention them and their support.
 - Personally, I tend not to thank people who were paid to do a specific job and only did that job. The acknowledgements is a section to name people who have gone out of their way to help you.
  - Your family for getting you to grad school (unless they really helped you do the work). However, in your thesis acknowledgements, you should absolutely include your family.

If in doubt it's probably a good idea to be as inclusive as you can in the acknowledgements.

>>>>

# The Acknowledgements {#acknowledgements}

@@

## Thank your funders

You should thank any people or organisations that funded you or any component of your work. Check with them they may have specific wording that they want you to use. Remember that your institution may not have given you money but may have supported you in other ways, and you should probably thank them. There are registers of funders, and you should use these to see whether your funder is present and exactly how they should be acknowledged. One example is the [Crossref Funder Registry](https://www.crossref.org/services/funder-registry/). Funders rely on being able to search [databases](#databases) for their name and find all of the work that they have funded. This means that you should use the particular wording that they supplied, especially if they specify this in their funding letter.

@@

## People and organisations to thank

 - Try to remember all of those people who helped you doing fieldwork, or in the laboratory. Especially laboratory technicians (if they are not also authors) without whom a lot of the work that we do isn’t possible.
 - People that may have helped administer your funds or other office administration. 
 - Organisations that gave you permission to do your work. In some journals, the acknowledgements are the correct place to put your permit numbers along with the names of the organisations who provided them. 
 - Increasingly, permission from ethics committees is placed in the materials and methods but do check with the journal instructions to authors.
 - People who supplied you with samples, images or other elements of your work or presentation of your manuscript.

@@

## Who not to thank (in your manuscript acknowledgements)
 - You do not need to thank anyone who is an author. The exception is in your thesis acknowledgements, where you should mention them and their support.
 - Personally, I tend not to thank people who were paid to do a specific job and only did that job. The acknowledgements is a section to name people who have gone out of their way to help you.
  - Your family for getting you to grad school (unless they really helped you do the work). However, in your thesis acknowledgements, you should absolutely include your family.

If in doubt, it's probably a good idea to be as inclusive as you can in the acknowledgements.
